
```{r}
Fancy = read.csv(file.choose())
Fancy.ts = ts(Fancy$Sales.Volume, start = c(1987,1),frequency = 12)
Fancy.ts
```
```{r}
plot(Fancy.ts)
```
```{r}
ggmonthplot(Fancy.ts)
```
```{r}
ggseasonplot(Fancy.ts)
```
```{r}
decompose(Fancy.ts)
```
```{r}
##Additive Method
plot(decompose(Fancy.ts))
```

```{r}
##Multiplicative Method
plot(decompose(Fancy.ts, type = "m"))
```

```{r}
hist(Fancy$Sales.Volume)
hist(decompose(Fancy.ts)$random, xlab = "Random",main = "Histogram of Additive Series")
```
```{r}
hist(decompose(Fancy.ts, type= "m")$random, xlab = "Random",main = "Histogram of Multiplicative Series")
```
##Spltting the data into Train and Test
```{r}
Fancy_Train = window(Fancy.ts, start = c(1987,1), end =c(1991,12), frequency = 12)
Fancy_Train
```

```{r}
Fancy_Test = window(Fancy.ts, start = c(1992,1), frequency=12)
Fancy_Test
```

```{r}
##Simple Exponential Smoothing
Fancy_Train_SES_Forecast = ses(Fancy_Train, h = 24)
summary(Fancy_Train_SES_Forecast)
```
```{r}
plot(Fancy_Train_SES_Forecast)
```

```{r}
Fancy_Test_SES = cbind(Fancy_Test, Fancy_Train_SES_Forecast$mean)
ts.plot(Fancy_Test_SES, col = c("blue", "red"), main = "Actual Vs Forecast-SES")
```

##Holt Model - Double Exponential
```{r}
Fancy_Train_Holt_Forecast = holt(Fancy_Train,h = 24)
summary(Fancy_Train_Holt_Forecast)
```
```{r}
plot(Fancy_Train_Holt_Forecast)
```
```{r}
Fancy_Test_Holt = cbind(Fancy_Test, Fancy_Train_Holt_Forecast$mean)
ts.plot(Fancy_Test_Holt, col = c("blue", "red"), main = "Actual Vs Forecast-Holt")
```
##Holt-Winters - Triple Exponential

```{r}
Fancy_Train_HW_Forecast = hw(Fancy_Train, h = 24)
summary(Fancy_Train_HW_Forecast)
```
```{r}
Fancy_Test_HW = cbind(Fancy_Test, Fancy_Train_HW_Forecast$mean)
Fancy_Test_HW_MAPE =mean(abs(Fancy_Test_HW[,1] - Fancy_Test_HW[,2])/Fancy_Test_HW[,1])
Fancy_Test_HW_MAPE
```
```{r}
Fancy_Test_HW_RMSE = sqrt(sum(Fancy_Test_HW[,1] - Fancy_Test_HW[,2])^2)
Fancy_Test_HW_RMSE
```

```{r}
ts.plot(Fancy_Test_HW, col = c("blue", "red"), main = "Actual Vs Forecast - HW Additive Series")
```
```{r}
Fancy_Train_HW_Forecast_M = hw(Fancy_Train, h = 24, seasonal = "m")
Train = cbind(Fancy_Train, Fancy_Train_HW_Forecast_M$fitted)
View(Train)
sqrt(mean((Train[,1] - Train[,2])^2))
summary(Fancy_Train_HW_Forecast_M)
```
```{r}
Fancy_Test_HW_M = cbind(Fancy_Test, Fancy_Train_HW_Forecast_M$mean)
Fancy_Test_HW_M_MAPE =mean(abs(Fancy_Test_HW_M[,1] - Fancy_Test_HW_M[,2])/Fancy_Test_HW_M[,1])
Fancy_Test_HW_M_MAPE
```
```{r}
View(Fancy_Test_HW_M)
sqrt(mean((Fancy_Test_HW_M[,1] - Fancy_Test_HW_M[,2])^2))
```

```{r}
ts.plot(Fancy_Test_HW_M, col = c("blue", "red"), main = "Actual Vs Forecast-HW-Multiplicative Series")

```
```{r}
Fancy_Train_HW_Forecast_M1 = hw(Fancy_Train, h = 24, seasonal = "m",alpha = 0.8, beta = 0.02, gamma = 0.0001)
summary(Fancy_Train_HW_Forecast_M1)
```
```{r}
Fancy_Test_HW_M1 = cbind(Fancy_Test, Fancy_Train_HW_Forecast_M1$mean)
Fancy_Test_HW_M1_MAPE =mean(abs(Fancy_Test_HW_M1[,1] - Fancy_Test_HW_M1[,2])/Fancy_Test_HW_M1[,1])
Fancy_Test_HW_M1_MAPE
```

```{r}
ts.plot(Fancy_Test_HW_M1, col = c("blue", "red"), main = "Actual Vs Forecast-HW-M1")
```

```{r}
Fancy_HW_5Years_Additive = hw(Fancy.ts,alpha = 0.4644,beta = 0.0001, gamma = 0.0002 ,h = 60)
summary(Fancy_HW_5Years_Additive)
```
```{r}
plot(Fancy_HW_5Years_Additive)
```

```{r}
Fancy_HW_5Years_M = hw(Fancy.ts,alpha = 0.2178,beta=0.0198, gamma = 0.0001,h = 60, seasonal = "m")
summary(Fancy_HW_5Years_M)
```
```{r}
plot(Fancy_HW_5Years_M)
```
```{r}
Fancy_HW_5Years_M1 = hw(Fancy.ts,alpha = 0.8,beta=0.02, gamma = 0.0001,h = 60, seasonal = "m")
summary(Fancy_HW_5Years_M1)
```
```{r}
plot(Fancy_HW_5Years_M1)
```

```{r}
Fancy_Train_STL = stl(Fancy_Train, s.window = 7)
Fancy_Train_STL_Forecast= forecast(Fancy_Train_STL,method = "rwdrift", h = 24)
summary(Fancy_Train_STL_Forecast)
```
```{r}
Fancy_Test_STL = cbind(Fancy_Test, Fancy_Train_STL_Forecast$mean)
View(Fancy_STL)
ts.plot(Fancy_Test_STL, col = c("blue", "red"), main = "Random Walk Drift Method on Train Data")

```
```{r}
Fancy.ts.STL = stl(Fancy.ts, s.window = 7)
Fancy.ts.STL.Forecast= forecast(Fancy.ts.STL,method = "rwdrift", h = 60)
summary(Fancy.ts.STL.Forecast)
```

```{r}
plot(Fancy.ts.STL.Forecast)
```

###Dicky-Fuller Test on Original TIme Series
```{r}
library(tseries)
adf.test(Fancy.ts)
```
```{r}
adf.test(diff(Fancy.ts))
```
```{r}
plot(diff(Fancy.ts), ylab = "First Order Difference")
```
```{r}
acf(Fancy.ts, lag = 60)
##The frequency is 12, hence every lag corresponds to 12 Months
```
```{r}
pacf(Fancy.ts, lag = 60)
```
```{r}
Fancy_Diff = diff(Fancy.ts)
acf(Fancy_Diff, lag = 60)
##There is an auto-correlation between the current year and sales 11 months ago for first-order differenced timeseries
```

```{r}
pacf(Fancy_Diff, lag = 60)
###Significant partial auto-correlation between the current year and sales 11 months ago

```
```{r}
Fancy.arima = arima(Fancy.ts, c(11,1,11))
Fancy.arima
```


```{r}
hist(Fancy.arima$residuals, col = "beige")
```
```{r}
Fancy.Arima.Fit = fitted(Fancy.arima)
ts.plot(Fancy.ts, Fancy.Arima.Fit, col = c("blue", "red"))
```
```{r}
Box.test(Fancy.arima$residuals, lag = 30, type = "Ljung-Box")
```

```{r}
auto.arima(Fancy.ts)
```
```{r}
ts.plot(Fancy.ts, fitted(auto.arima(Fancy.ts)),col = c("blue", "red") )
```

```{r}
Box.test(auto.arima(Fancy.ts)$residuals, lag = 30, type = "Ljung-Box")
```
```{r}
adf.test(Fancy_Train)
```
```{r}
adf.test(diff(Fancy_Train))
```
```{r}
acf(diff(Fancy_Train))
```
```{r}
pacf(diff(Fancy_Train))
```
```{r}
Fancy_Train_Arima = arima(Fancy_Train, c(12,1,12))
summary(Fancy_Train_Arima)
```
```{r}
Fancy.Arima.Train.Fit = fitted(Fancy_Train_Arima)
ts.plot(Fancy_Train, Fancy.Arima.Train.Fit, col = c("blue", "red"), main = "Actual Train Data Vs Fitted Train Data")
```
```{r}
Box.test(Fancy_Train_Arima$residuals, lag = 30, type = "Ljung-Box")
```

```{r}
Fancy_Train_Arima_Forecast = forecast(Fancy_Train_Arima, h=24)
Fancy_Test_Arima = cbind(Fancy_Test,Fancy_Train_Arima_Forecast$mean)
View(Fancy_Test_Arima)
ts.plot(Fancy_Test_Arima, col = c("blue", "red"), main = "Arima Model applied on Test Data")
```
```{r}
forecast(arima(Fancy.ts, c(12,1,12)), h = 60)
plot(forecast(arima(Fancy.ts, c(12,1,12)), h = 60), main = "Forecast for 5 Years based on ARIMA Model")
```
##Auto-Arima
```{r}
auto.arima(Fancy_Train)
```
```{r}
Fancy_Train_Autoarima = forecast(auto.arima(Fancy_Train), h = 24)
```
```{r}
Fancy_Test_Autoarima = cbind(Fancy_Test, Fancy_Train_Autoarima$mean)
ts.plot(Fancy_Test_Autoarima, col = c("blue", "red"))
```
```{r}
summary(Fancy_Train_Autoarima)
```
```{r}
Box.test(auto.arima(Fancy_Train)$residuals, lag = 30, type = "Ljung-Box")
```
###SARIMA
```{r}
Fancy_Train_ArimaSeasonal = arima(Fancy_Train, order = c(2,1,2), seasonal = list(order = c(1,1,2), period = 12))
```

```{r}
Fancy_Train_ArimaSeasonal
```
```{r}
library(forecast)
Fancy_Train_ArimaSeasonal_Forecast = forecast(Fancy_Train_ArimaSeasonal, h = 24)
summary(Fancy_Train_ArimaSeasonal_Forecast)
```

```{r}
Fancy_Test_ArimaSeasonal = cbind(Fancy_Test, Fancy_Train_ArimaSeasonal_Forecast$mean)
ts.plot(Fancy_Test_ArimaSeasonal, col = c("blue", "red"), main = "ARIMA(2,1,2)(1,1,2)[12] Model Test Data Validation using Forecast func.")
```
```{r}
Box.test(Fancy_Train_ArimaSeasonal$residuals, lag = 30, type = "Ljung-Box")
```
```{r}
forecast()
```


```{r}
library(forecast)
ntest = Arima(Fancy_Test, model=Fancy_Train_ArimaSeasonal, include.drift = FALSE, include.mean = FALS)
Fancy_ntest = cbind(Fancy_Test, ntest$fitted)
ts.plot(Fancy_ntest, col = c("blue", "red"), main = "ARIMA(2,1,2)(1,1,2)[12] Model Test Data Validation using Arima func.")

```
```{r}
SArima = forecast(Arima(Fancy.ts, model=Fancy_Train_ArimaSeasonal), h = 60)
```


```{r}
plot(forecast(Arima(Fancy.ts, model=Fancy_Train_ArimaSeasonal), h = 60))

```


###Combined
```{r}
Fancy_Test_Combined = cbind(Fancy_Test,
Fancy_Train_HW_Forecast_M$mean,
Fancy_Train_HW_Forecast_M1$mean,
Fancy_Train_STL_Forecast$mean,
Fancy_Train_ArimaSeasonal_Forecast$mean, ntest$fitted)

colnames(Fancy_Test_Combined) = c("TestData","HW-M","HW-M1","STL","Sarima", "SArima")
View(Fancy_Test_Combined)

View(Fancy_Test_Combined)
getwd()
write.csv(Fancy_Test_Combined, "Summary.csv")
ts.plot(Fancy_Test_Combined, col = c("blue","violet","orange","black", "red", "yellow"))
```
```{r}
i = 48
pred_ets = c()
pred_arima = c()
while(i <= nrow(Fancy)){
  ts <- ts(Fancy[1:i, Fancy$Sales.Volume], start=c(1987, 1), frequency=12)

  pred_ets <- rbind(pred_ets, data.frame(forecast(ets(ts), 3)$mean[1:3]))
	  pred_arima <- rbind(pred_arima, data.frame(forecast(auto.arima(ts), 3)$mean[1:3]))

  i = i + 3
}
```






A few line of codes below are on ARIMA which is a part of the report as outputs, but the data for these codes are a scaled version of the original data.(Attached the scaled version of the data too along with the report submission)





```{r}

sales <- read.csv(file.choose())

sales <- ts(sales[,4], start=c(1987,1), frequency=12)

plot(sales)
```


```{r}
# train test split of 70- 30 
salesTrain <- window(sales, start = c(1987,1), end = c(1991,12), frequency = 12)
salesTest <- window(sales, start = c(1992,1), frequency=12)

salesTrain
salesTest
plot(salesTrain)
plot(salesTest)
```


```{r}

library(dplyr)

arima.fit <- arima(salesTrain, c(2,1,1))
arima.fit
summary(arima.fit)

data.fit <- fitted(arima.fit)
ts.plot(salesTrain, data.fit, col= c("blue", "red"), main = "ARIMA(2,1,1) Model: Train vs Fitted Values")





arima.fit %>% forecast( h = 24) %>% autoplot() + autolayer(salesTest)
#summary(forecast(arima.fit,h = 24))
#plot(forecast(arima.fit,h = 24))

# on test data
ntest <- Arima(salesTest, model = arima.fit)
ts.plot(salesTest, ntest$fitted, col= c("blue","red"))
accuracy(ntest)
ntest$fitted

summary(ntest)

ntest$fitted

arima.fl <- (arima(sales, c(2,1,1)))

arima.fl %>%forecast(h = 60)%>%autoplot
```

analysing residuals 

```{r}
hist(arima.fit$residuals, col = "beige")
hist(arima.fit.test$residuals, col = "beige")

```

Box test 
```{r}
acf(arima.fit$residuals)
acf(arima.fit.test$residuals)

# box test - portmanteau  test

Box.test(arima.fit$residuals, lag =5, type = "Ljung-Box")
Box.test(arima.fit.test$residuals, lag = 6, type = "Ljung-Box")

```


```{r}
auto.arima(salesTrain)
```

Arima Model with seasonality

```{r}

library(forecast)
arima.fit.s <- arima(salesTrain, order = c(2,1,2), seasonal = list(order = c(1,1,2), period = 12))
arima.fit.s
accuracy(arima.fit.s)
summary(arima.fit.s)

data.fit.s <- fitted(arima.fit.s)
ts.plot(salesTrain, data.fit.s, col= c("blue", "red"), main = "ARIMA(2,1,2)(1,1,2)[12] Train Vs Fitted Values ")

arima.fit.s %>% forecast( h = 24) %>% autoplot() + autolayer(salesTest)
#summary(forecast(arima.fit,h = 24))
#plot(forecast(arima.fit,h = 24))

# on test data
ntest.s <- Arima(salesTest, model = arima.fit.s)
ts.plot(salesTest, ntest.s$fitted, col = c("blue", "red"))
ts.plot(salesTest, ntest.s$fitted, col= c("blue","red"), main = "ARIMA(2,1,2)(1,1,2)[12] Test Vs Fitted Values ")

ts.plot(salesTest, col = c("blue"))
ts.plot(ntest.s$fitted, col= c("red"))
accuracy(ntest.s)


plot(forecast(arima.fit.s, h = 24))


# box test 
Box.test(arima.fit.s$residuals, lag =5, type = "Ljung-Box")
Box.test(ntest.s$residuals, lag = 6, type = "Ljung-Box")

# forecast on entire data

arima.fit.sf <- arima(sales, order = c(2,1,2), seasonal = list(order = c(1,1,2), period = 12))

summary(arima.fit.sf)
  plot(forecast(arima.fit.sf, h = 60),main = " Forecast for 5 Years using ARIMA(2,1,2)(1,1,2)[12]")

forecast(arima.fit.sf, h = 60)
```


